Comb associative operads

### 2018-07-09 CAs_operads.tex
Partie 1 chantier (suite)
Remarques de Cyrille sur partie 1

In order to show that the poset $\left(\CAsAll, \OrdCAs\right)$ is also
a lattice, that is two elements of $\CAsAll$ admit lower-bounds and
upper-bounds, we relate $\left(\CAsAll, \OrdCAs\right)$ with the lattice
of integers $\left(\N,\mid, \gcd, \lcm\right)$, where $\mid$ denotes the
division relation, $\gcd$ denotes the greatest common divisor, and
$\lcm$ the least common multiple operators, respectively. In order to
obtain this relationship, we introduce the notion of \Def{left rank} for
any binary tree $\Tfr$ as the number $\LBranch(\Tfr)$ of internal nodes
in the left branch beginning at the root of $\Tfr$.
For instance,
\begin{equation}
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \xmapsto{\; \LBranch\; } 2
    \qquad \mbox{and} \qquad
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \xmapsto{\; \LBranch\; } 2.
\end{equation}
The left rank of $\Tfr$ is represented as follows: \Todo{???}
\medbreak

\begin{Lemma}\label{lem:left_rank_and_CongrCAs}
    Let $\gamma$ be a strictly positive integer and let $\Tfr$ and
    $\Tfr'$ be two binary trees. If $\Tfr \CongrCAs{\gamma} \Tfr'$, then
    \begin{equation}
        \LBranch(\Tfr) \mod \gamma - 1
        \enspace = \enspace
        \LBranch(\Tfr') \mod \gamma - 1.
    \end{equation}
\end{Lemma}
\begin{proof}
    Let $\Rew$ be the rewrite relation on $\Mag$ induced by the rewrite
    rule $\LComb{\gamma} \Rew \RComb{\gamma}$. In the rest of the
    proof, $\Tfr$ and $\Tfr$ are two binary trees.
    \smallbreak

    First, we show that $\Tfr \Rew \Tfr'$ implies that
    $\LBranch\left(\Tfr'\right) - \LBranch\left(\Tfr\right)$ is
    divisible by $\gamma - 1$. \Todo{à faire.}
    \smallbreak

    Assume that $\Tfr \RewRT \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in \{1, \dots, k - 1\}$, we have
    $\Tfr_i\Rew \Tfr_{i + 1}$. The integer
    \begin{math}
        \LBranch\left(\Tfr_{i + 1}\right)
        -
        \LBranch\left(\Tfr_i\right)
    \end{math}
    is divisible by $\gamma - 1$ from the first part of the proof, so
    that $\LBranch\left(\Tfr'\right) - \LBranch\left(\Tfr\right)$ is
    divisible by $\gamma-1$.
    \smallbreak

    Assume that $\Tfr \RewRTS \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in \{1, \dots, k - 1\}$, we have
    $\Tfr_i \RewRTS \Tfr_{i + 1}$ or $\Tfr_{i + 1} \RewRTS \Tfr_i$.
    From the second part of the proof, the integer
    \begin{math}
        \LBranch\left(\Tfr_{i + 1}\right)
        -
        \LBranch\left(\Tfr_i\right)
    \end{math}
    is divisible by $\gamma - 1$, so that
    $\LBranch\left(\Tfr'\right) - \LBranch\left(\Tfr\right)$ is
    divisible by $\gamma - 1$.
    \smallbreak

    The congruence $\CongrCAs{\gamma}$ is the equivalence relation
    induced by $\Rew$, that is we have $\Tfr \CongrCAs{\gamma} \Tfr'$ if
    and only if $\Tfr \RewRTS \Tfr'$. Hence, from the third part of
    the proof, $\Tfr \CongrCAs{\gamma} \Tfr'$ implies that
    \begin{math}
        \LBranch\left(\Tfr'\right) - \LBranch\left(\Tfr\right)
    \end{math}
    is divisible by $\gamma - 1$.
\end{proof}
\medbreak

\begin{Proposition} \label{prop:division_CAs}
    Let $\gamma$ and $\gamma'$ be two positive integers.Then, there
    exists a morphism $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ if and
    only if $\left(\gamma - 1\right) \mid \left(\gamma' - 1\right)$.
\end{Proposition}
\begin{proof}
    From Lemma~\ref{lem:morphism_CAs}, it is enough to show that
    $\LComb{\gamma'} \CongrCAs{\gamma} \RComb{\gamma'}$ if and only if
    $\left(\gamma-1\right) \mid \left(\gamma'-1\right)$. If
    $\LComb{\gamma'}\CongrCAs{\gamma}\RComb{\gamma'}$, from
    Lemma~\ref{lem:left_rank_and_CongrCAs},
    \begin{math}
        \LBranch\left(\LComb{\gamma'}\right)
        - \LBranch\left(\RComb{\gamma'}\right)
        = \gamma' - 1
    \end{math}
    is divisible by $\gamma-1$, which shows the direct implication.
    \smallbreak

>>>>

In order to show that the poset $\left(\CAsAll, \OrdCAs\right)$ is also
a lattice, that is two elements of $\CAsAll$ admit lower-bounds and
upper-bounds, we relate $\left(\CAsAll, \OrdCAs\right)$ with the lattice
of integers $\left(\N,\mid, \gcd, \lcm\right)$, where $\mid$ denotes the
division relation, $\gcd$ denotes the greatest common divisor, and
$\lcm$ the least common multiple operators, respectively.
\medbreak

\begin{Lemma}\label{lem:left_rank_and_CongrCAs}
    Let $\gamma$ be a strictly positive integer and let $\Tfr$ and
    $\Tfr'$ be two binary trees. If $\Tfr \CongrCAs{\gamma} \Tfr'$, then
    \begin{equation}
        \LRank(\Tfr) \mod \gamma - 1
        \enspace = \enspace
        \LRank(\Tfr') \mod \gamma - 1.
    \end{equation}
\end{Lemma}
\begin{proof}
    Let $\Rew$ be the rewrite relation on $\Mag$ induced by the rewrite
    rule $\LComb{\gamma} \Rew \RComb{\gamma}$. In the rest of the
    proof, $\Tfr$ and $\Tfr$ are two binary trees.
    \smallbreak

    First, we show that $\Tfr \Rew \Tfr'$ implies that
    $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by $\gamma - 1$. \Todo{à faire.}
    \smallbreak

    Assume that $\Tfr \RewRT \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in \{1, \dots, k - 1\}$, we have
    $\Tfr_i\Rew \Tfr_{i + 1}$. The integer
    \begin{math}
        \LRank\left(\Tfr_{i + 1}\right)
        -
        \LRank\left(\Tfr_i\right)
    \end{math}
    is divisible by $\gamma - 1$ from the first part of the proof, so
    that $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by $\gamma-1$.
    \smallbreak

    Assume that $\Tfr \RewRTS \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in \{1, \dots, k - 1\}$, we have
    $\Tfr_i \RewRTS \Tfr_{i + 1}$ or $\Tfr_{i + 1} \RewRTS \Tfr_i$.
    From the second part of the proof, the integer
    \begin{math}
        \LRank\left(\Tfr_{i + 1}\right)
        -
        \LRank\left(\Tfr_i\right)
    \end{math}
    is divisible by $\gamma - 1$, so that
    $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by $\gamma - 1$.
    \smallbreak

    The congruence $\CongrCAs{\gamma}$ is the equivalence relation
    induced by $\Rew$, that is we have $\Tfr \CongrCAs{\gamma} \Tfr'$ if
    and only if $\Tfr \RewRTS \Tfr'$. Hence, from the third part of
    the proof, $\Tfr \CongrCAs{\gamma} \Tfr'$ implies that
    \begin{math}
        \LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)
    \end{math}
    is divisible by $\gamma - 1$.
\end{proof}
\medbreak

\begin{Proposition} \label{prop:division_CAs}
    Let $\gamma$ and $\gamma'$ be two positive integers.Then, there
    exists a morphism $\varphi:\CAs{\gamma'} \to \CAs{\gamma}$ if and
    only if $\left(\gamma - 1\right) \mid \left(\gamma' - 1\right)$.
\end{Proposition}
\begin{proof}
    From Lemma~\ref{lem:morphism_CAs}, it is enough to show that
    $\LComb{\gamma'} \CongrCAs{\gamma} \RComb{\gamma'}$ if and only if
    $\left(\gamma-1\right) \mid \left(\gamma'-1\right)$. If
    $\LComb{\gamma'}\CongrCAs{\gamma}\RComb{\gamma'}$, from
    Lemma~\ref{lem:left_rank_and_CongrCAs},
    \begin{math}
        \LRank\left(\LComb{\gamma'}\right)
        - \LRank\left(\RComb{\gamma'}\right)
        = \gamma' - 1
    \end{math}
    is divisible by $\gamma-1$, which shows the direct implication.
    \smallbreak

### 2018-07-09 Operads_rewrite.tex
Remarques de Cyrille sur partie 1

Let us provide some elementary definitions and notations about operads.
If $x$ is an element of $\Oca$ such that $x \in \Oca(n)$ for an
$n \geq 1$, the \Def{arity} $|x|$ of $x$ is $n$. The
\Def{complete composition maps} of $\Oca$ are the map
\begin{equation}
    \circ : \Oca(n) \times
    \Oca\left(m_1\right) \times \dots \times \Oca\left(m_n\right)
    \to \Oca\left(m_1 + \dots + m_n\right)
\end{equation}
defined, for any $x \in \Oca(n)$ and $y_1, \dots, y_n \in \Oca$, by
\begin{equation}
    x \circ \left[y_1, \dots, y_n\right] :=
    \left(\dots \left(\left(x \circ_n y_n\right)
    \circ_{n - 1} y_{n - 1}\right) \dots\right) \circ_1 y_1.
\end{equation}
If $\Oca_1$ and $\Oca_2$ are two operads, a map
$\phi : \Oca_1 \to \Oca_2$ is an \Def{operad morphism} if it respects
the arities, sends the unit of $\Oca_1$ to the unit of $\Oca_2$, and
commutes with partial composition maps. We say that $\Oca_2$ is a
\Def{suboperad} of $\Oca_1$ if $\Oca_2$ is a subset of $\Oca_1$
containing the unit of $\Oca_1$ and the partial composition maps of
$\Oca_2$ are the ones of $\Oca_1$ restricted on $\Oca_2$. For any subset
$\GeneratingSet$ of $\Oca$, the \Def{operad generated} by
$\GeneratingSet$ is the smallest suboperad $\Oca^\GeneratingSet$ of
$\Oca$ containing $\GeneratingSet$. When $\Oca = \Oca^\GeneratingSet$,
we say that $\GeneratingSet$ is a \Def{generating set} of $\Oca$. In
this case, and when additionally $\GeneratingSet$ is minimal among all
the subsets of $\Oca$ satisfying this property, $\GeneratingSet$ is a
\Def{minimal} generating set of $\Oca$. An \Def{operad congruence} is an
equivalence relation $\Congr$ on $\Oca$ such that $\Congr$ respects the
arities, and for any $x, x', y, y' \in \Oca$ such that $x \Congr x'$ and
$y \Congr y'$, $x \circ_i y$ is $\Congr$-equivalent to $x' \circ_i y'$
for any valid integer $i$. Given an operad congruence $\Congr$, the
\Def{quotient operad} $\Oca/_{\Congr}$ of $\Oca$ by $\Congr$ is the
operad of all $\Congr$-equivalence classes endowed with the partial
composition maps defined in the obvious way. In the case where all the
sets $\Oca(n)$, $n \geq 1$, are finite, the \Def{Hilbert series}
$\HilbertSeries_\Oca(t)$ of $\Oca$ is the series defined by
\begin{equation}
    \HilbertSeries_\Oca(t) := \sum_{n \geq 1} \# \Oca(n) \; t^n.
\end{equation}
\medbreak

@@

\Def{left rank} of $\Tfr$ as the number $\LBranch(\Tfr)$ of internal

@@

    \xmapsto{\; \LBranch\; } 3.

@@

Equivalently, $\LBranch(\Tfr)$ is the length of the prefix of

@@

satisfying

@@

        \draw[Edge,ultra thick](1)--(5);

@@

        \draw[Edge,ultra thick](3)--(1);

@@

        \draw[Edge,ultra thick](5)--(3);

@@

@@

@@

        \draw[Edge,ultra thick](11)--(9);

@@

        \draw[Edge,ultra thick](9)--(17);

@@

        \draw[Edge,ultra thick](17)--(11);

@@

>>>>

Let us provide some elementary definitions and notations about operads.
If $x$ is an element of $\Oca$ such that $x \in \Oca(n)$ for an
$n \geq 1$, the \Def{arity} $|x|$ of $x$ is $n$. The
\Def{complete composition maps} of $\Oca$ are the map
\begin{equation}
    \circ : \Oca(n) \times
    \Oca\left(m_1\right) \times \dots \times \Oca\left(m_n\right)
    \to \Oca\left(m_1 + \dots + m_n\right)
\end{equation}
defined, for any $x \in \Oca(n)$ and $y_1, \dots, y_n \in \Oca$, by
\begin{equation}
    x \circ \left[y_1, \dots, y_n\right] :=
    \left(\dots \left(\left(x \circ_n y_n\right)
    \circ_{n - 1} y_{n - 1}\right) \dots\right) \circ_1 y_1.
\end{equation}
If $\Oca_1$ and $\Oca_2$ are two operads, a map
$\phi : \Oca_1 \to \Oca_2$ is an \Def{operad morphism} if it respects
the arities, sends the unit of $\Oca_1$ to the unit of $\Oca_2$, and
commutes with partial composition maps. We say that $\Oca_2$ is a
\Def{suboperad} of $\Oca_1$ if $\Oca_2$ is a subset of $\Oca_1$
containing the unit of $\Oca_1$ and the partial composition maps of
$\Oca_2$ are the ones of $\Oca_1$ restricted on $\Oca_2$. For any subset
$\GeneratingSet$ of $\Oca$, the \Def{operad generated} by
$\GeneratingSet$ is the smallest suboperad $\Oca^\GeneratingSet$ of
$\Oca$ containing $\GeneratingSet$. When $\Oca = \Oca^\GeneratingSet$,
we say that $\GeneratingSet$ is a \Def{generating set} of $\Oca$. In
this case, and when additionally $\GeneratingSet$ is minimal among all
the subsets of $\Oca$ satisfying this property, $\GeneratingSet$ is a
\Def{minimal} generating set of $\Oca$. An \Def{operad congruence} is an
equivalence relation $\Congr$ on $\Oca$ such that $\Congr$ respects the
arities, and for any $x, x', y, y' \in \Oca$ such that $x \Congr x'$ and
$y \Congr y'$, $x \circ_i y$ is $\Congr$-equivalent to $x' \circ_i y'$
for any valid integer $i$. Given an operad congruence $\Congr$, the
\Def{quotient operad} $\Oca/_{\Congr}$ of $\Oca$ by $\Congr$ is the
operad of all $\Congr$-equivalence classes endowed with the partial
composition maps defined in the obvious way. In the case where all the
sets $\Oca(n)$, $n \geq 1$, are finite, the \Def{Hilbert series}
$\HilbertSeries_\Oca(t)$ of $\Oca$ is the series defined by
\begin{equation} \label{equ:hilbert_series}
    \HilbertSeries_\Oca(t) := \sum_{n \geq 1} \# \Oca(n) \; t^n.
\end{equation}
\medbreak

We have provided here definitions about operads in the category of sets.
Nevertheless, operads can be defined in the category of $\K$-vector
spaces, where $\K$ denotes in all this text a field of characteristic
zero. We call them \Def{linear operads}. All the above definitions
extend for linear operads, mainly by substituting Cartesian products
$\times$ of sets with tensor products $\otimes$ of spaces, maps with
linear maps, operad congruences with operad ideals, and cardinalities
of sets with space dimensions (for instance
in~\eqref{equ:hilbert_series}). Besides, when $\Oca$ is a linear operad
admitting a basis $B$ such that its unit $\Unit$ belongs to $B$ and all
partial composition maps are internal in $B$, $\Oca$ can be studied as a
set-theoretic operad.
\medbreak

@@

\Def{left rank} of $\Tfr$ as the number $\LRank(\Tfr)$ of internal

@@

    \xmapsto{\; \LRank\; } 3.

@@

Equivalently, $\LRank(\Tfr)$ is the length of the prefix of

@@

containing the single rewrite rule
\begin{equation} \label{equ:example_rewrite_rule}

@@

        \draw[EdgeColorF](1)--(5);

@@

        \draw[EdgeColorF](3)--(1);

@@

        \draw[EdgeColorF](5)--(3);

@@

    \end{tikzpicture}\,,

@@

\begin{equation} \label{equ:example_rewrite_step}

@@

        \draw[EdgeColorF](11)--(9);

@@

        \draw[EdgeColorF](9)--(17);

@@

        \draw[EdgeColorF](17)--(11);

@@

The right member of~\eqref{equ:example_rewrite_step} is obtained by
replacing, in the tree of left member
of~\eqref{equ:example_rewrite_step}, a subtree equal to the left
member of~\eqref{equ:example_rewrite_rule} starting at the right child
of its root by the right member of~\eqref{equ:example_rewrite_rule}.
